### 2004-07-13 Documentation/part2.tex
added section about globes put in the fdl part on page two, made the sections of fdl disappear from toc added multiple index support changed the \GLB and \GLBNS commands such, that they prepend 'glb' in the text and add their argument to the api index. makeindex has not be called four times: 	makeindex norm.idx 	makeindex api.idx 	makeindex aedl.idx 	makeindex constants.idx
Adding entries now works with \index{index_name}{index_entry}, e.g.
\index{norm}{Blau} adds 'Blau' to the usual index.

In order to implement a experiment definition including various
sources of systematical errors, we use several abstraction levels. 
The first level is the so-called ``channel'', which is the link between 
the oscillation physics and the detection properties for a specfific oscillation pattern (\cf, \figu{channel}). A channel specifies the mapping of a specific neutrino flavor produced by the source onto a reconstructed neutrino flavor.
For example, a muon neutrino oscillates into an electron neutrino and subsequently interacts via quasi-elastic charged current scattering. The measured energy and direction of
the secondary electron in the detector then allows to reconstruct the neutrino energy. The connection from the source flux of the muon neutrino, via the  probability to appear as a electron neutrino, to its detection properties (such as cross sections and energy smearing) is encapsulated into the channel.

@@

The channels are the building blocks for the so-called ``rules''. In general, a rule consists of one or more ``signal'' and ``background'' oscillation channels, which are normalized with efficiencies
(\cf, \figu{rule}). The event numbers from these channels are added {\em before} the $\Delta \chi^2$-value is calculated.\footnote{Note that
in this manual, we $\chi^2$ and $\Delta \chi^2$ are equal, since
for simulated data $\Delta \chi^2 = 0$ at the best-fit point. Thus, we
are using $\chi^2$ and $\Delta \chi^2$ as equal quantities.} In addition, each rule implements an {\em independent} systematics, such as signal and background normalization errors. Eventually, each rule gives a $\Delta \chi^2$-value, and the total $\Delta \chi^2$ of one experiment is obtained by adding the $\Delta \chi^2$'s of all rules (\cf, \figu{experiment}). 
 An example for a rule could look like this: We want to detect electron
  neutrino appearance (``signal''), where the overall efficiency for 
  quasi-elastics electron neutrino events is $0.4$. There is a fraction of 
  $0.01$ of all neutral current events which are mis-identified as 
  quasi-elastic electron neutrino events (``background''). The neutral 
  current fraction is only known within $10\%$ (``background uncertainty'') 
  and there is
an energy scale uncertainty of $100\,\mathrm{MeV}$ (``energy calibration error'').
All this systematics is independent of the other rules.  Thus, a rule connects the event rates to the calculation of a $\Delta \chi^2$ which properly includes systematical errors. The resulting $\Delta \chi^2$ is then the starting point for the oscillation physics analysis. Note again that
\begin{itemize}
\item
 Within each rule the event numbers are added.
\item
 Within each rule the systematics is treated independently from the other rules.
\item
 For each rule the $\Delta \chi^2$ is computed; the $\Delta \chi^2$'s from all rules are added.
\end{itemize}

@@

If you want to have several experiments in one file, separate the different
 experiments by 
\begin{quote}
{\tt    \#NEXT\#}
\end{quote}
This command resets the counters for channels, rules and energy resolution
functions. All variables have their scope limited 
by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}.  This allows 
a consistent treatment of various experiments in one file.

As another feature of AEDL one can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
 for nested includes, too. The included file is not required to begin 
 with {\tt \%!GLoBES} to facilitate cut and paste:
\begin{quote}
{\tt include "./file\_1"}
\end{quote}
With this include mechanism, one can use constructions such as 
\begin{quote}
{\tt    include  "NuFact.gls"\\
        \#NEXT\#\\
        include   "JHFHK.gls"
}
\end{quote}
in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. As one sees from the example above
file names are always quoted in \AEDL.\index{File names}
Even if one uses the
automatic variable {\tt \#CC} in both experiments, 
but the cross section data are different (for example, because of different target nuclei), the correct 
cross section data will be applied to each of the experiments. 
Note that, alternatively, one can 
also load both files successively.

Furthermore, one can define constants such as
\begin{quote}
{\tt
Pi = 3.141
}
\end{quote}
These constants can not only be defined within one \AEDL\ file but also
by the calling C program, which allows to use a simple but powerful variable
substitution mechanism as described in~\Sec~\ref{sec:aedlparams}.
\index{External \AEDL\ parameters}

Alos some simple algebraic manipulations are possible, such as
\begin{quote}
{\tt
Pi+1\\
\verb+Pi^2+\\
sin(Pi/2)\\
}
\end{quote}
The following mathematical functions from {\tt <math.h>} are available: 
{\tt sin}, {\tt cos}, {\tt tan}, {\tt asin}, {\tt acos}, {\tt atan}, 
{\tt log}, {\tt log10}, {\tt exp}, {\tt sqrt}. 
These functions can be used everywhere, where
otherwise only a scalar number would appear. However, they can not be
applied to lists, such as {\tt sin(\{1,2,3\})} will not work.

@@

The quantity which can be used to scale the overall integrated luminosity of an experiment, is the fiducial detector mass. For example,
\begin{quote}
{\tt \$target\_mass = 50.0 }
\end{quote}
defines a $50 \, \mathrm{kt}$ detector for a neutrino factory.

There are two principal ways to initialize a neutrino flux: 
Either one can use 
a built-in source, or one can provide a file. In both cases,
a flux is defined by the environment {\tt flux}, such as
\begin{quote}
  {\tt flux(\#name)<\\
\tb $\ldots$\\
\tb @time = 8.0 \\
>}
\end{quote}
with a running time of $8$ years. Note that the running time is used within
the {\tt flux} environment. This feature can be used to load the neutrino and antineutrino fluxes separately, in order to combine them with different
running times within one experiment. The name of the flux {\tt \#name} will later be refered to in the channel definitions.

For a built-in neutrino source, one has to specify which
built-in spectrum has to be used, as well as its parameters. The software
will then automatically calculate the neutrino spectrum. Note that in this
case, there is no degree of freedom in the choice of the source units.
Currently, two built-in fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}). In these cases, the muon energy (enery of the parent particle) has to be specified together with the number of useful decays  per year. Thus, an example to set up a neutrino factory flux is
\begin{quote}
{\tt flux(\#mu\_plus)<\\
\tb  @builtin = 1\\
\tb  @parent\_energy = 50.0\\
\tb  @stored\_muons = 5.33e+20\\
\tb  @time = 8.0\\
>}
\end{quote}
%
For a user-defined flux, one has to give it the file name:
\begin{quote}
{\tt flux(\#user)<}\\
{\tt \tb @flux\_file = "user\_file\_1.dat"\\
\tb @time = 2.0\\
\tb @power = 4.0\\
\tb @norm = 1e+8}\\
{\tt >}
\end{quote}
In this case, the {\tt @norm} variable is an overall normalization which defines a conversion factor from the fluxes in the file to the units in \GLOBES . In general, there are many ways to give the source power of a 
neutrino source, such as neutrinos per proton on target per area per time frame. Right now, each flux has its own normalization factor, which is
not always straightfoward to calculate. Often, one has to take into account
many things, such as the number of target particles per unit mass. 
In addition, the fluxes will be rescaled by $1/L^2$, which means that the
normalization must contain a factor $L_0^2$. Here $L_0$ is the distance from the source for which the flux is given to the actual neutrino production region. At the end, it is left to the user to ensure that the 
numbers in the flux file give, after the multiplication with {\tt @norm}, 
the proper numbers of produced neutrinos corresponding to the chosen target power {\tt @power}. Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.

The software assumes that the given flux file has seven columns and
501 lines with equidistant energies. The format is:
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
In order to access fluxes at arbitrary energies, linear interpolation 
is used by the software. In general, it is advisable to provide the flux between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}), since these values are used by the software. However, if the energy leaves the range of values given in the file, zero is returned. 
Note that unused (and therefore often unknown) fluxes can be initialized with zeros, but can not just be omitted. For example, if one wants to define an experiment with different running times with neutrinos and antineutrinos, one can just use the same flux file in two different {\tt flux} environments with two different running times {\tt @time}. The unused fluxes in each case will then just be ignored.

@@

The baseline\index{Baseline} is given by
\begin{quote}
{\tt \$baseline = 3000.0 }
\end{quote}
Note that baseline lengths are always assumed to be in
kilometers.

@@

Furthermore, the matter density profile along the baseline
has to be specified. The simplest matter density profile is a constant matter density profile with the average matter density from the PREM\cite{Stacey} onion shell model of the Earth\index{PREM}\index{Earth matter density}:
\begin{quote}
{\tt \$profiletype=1 }
\end{quote}
%
For a better approximation of the realistic Earth matter density profile, one can use an arbitrary number of equidistant steps of the PREM profile:
\begin{quote}
{\tt \$profiletype=2 } \\
{\tt \$densitysteps=20 }
\end{quote}
Note that the value of {\tt \$densitysteps} is time-critical, since the
computation time of oscillation probabilities is directly 
proportional to the number of layers.
%
As a  third possibility, one can specify the matter density profile 
manually with a list of thicknesses and densities of the matter density layers. This example uses two density steps with two different densities:
\begin{quote}
{\tt \$profiletype=3 } \\
{\tt \$densitytab=\{2.8, 3.5\}}\\
{\tt \$lengthtab=\{1000.0, 2000.0\}}\\
\end{quote}
It is important that both lists habe the same length and that the  thicknesses given in  {\tt \$lengthtab} add up to the value of
{\tt \$baseline}. In addition, matter densities are always given in $g/cm^3$.
%
This approach can also be used for a constant matter density profile with
a specific matter density:
\begin{quote}
{\tt \$profiletype=3 } \\
{\tt \$densitytab=\{3.5\}}\\
{\tt \$lengthtab=\{3000.0\}}\\
\end{quote}
The possible options for matter density profiles are summarized in \Tab~\ref{tab:profiletypes}.

@@

Cross sections\index{Cross sections} will later be used as part of the 
channel definition (see \Sec~\ref{sec:channel}). Similar to the source 
fluxes, they are provided by the user as a data file:
\begin{quote}
{\tt cross(\#name)<}\\
{\tt \tb @cross\_file ="user\_file\_1.dat"}\\
{\tt >}
\end{quote}  
This cross section can later be refered to by {\tt \#name}.

Cross sections are in \GLOBES\ given as differential cross section per energy:
\begin{equation}
\hat\sigma(E)=\sigma(E)/E\,\left[ 10^{-38}\,
\frac{\mathrm{cm}^2}{\mathrm{GeV}^2} \right]
\end{equation}
The software assumes that the cross section files are text files with 
seven columns and $1001$ lines of the form
\begin{quotation}
$\mathrm{log}_{10} E\quad
\hat\sigma_{\nu_e}\quad
\hat\sigma_{\nu_\mu}\quad
\hat\sigma_{\nu_\tau}\quad
\hat\sigma_{\bar\nu_e}\quad
\hat\sigma_{\bar\nu_\mu}\quad
\hat\sigma_{\bar\nu_\tau}$
\end{quotation}
Here the logarithms of the energy values have to be equidistant. For 
arbitrary energies, linear interpolation is used. If the energy leaves the
range of values given in the file, $0.0$ will be assumed. In general, it is advisable to provide the cross sections in the range between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}).
Unused cross sections have to be filled with zeros, and can not
 be just omitted.

@@

Channels\index{Channel} in \GLOBES\ represent an intermediate level 
between the pure oscillation physics given by the oscillation probability
$P_{\alpha\beta}$, and the total event rates composed of signal and 
background. A channel describes the path from one
initial neutrino flavor in the source to the event rates in the detector for one specific interaction type (IT) and final flavor.  
Therefore, a channel contains the description of the 
initial neutrino flavor, its CP eigenvalue (neutrino or antineutrino), 
the detected neutrino flavor, the interaction cross sections for the chosen interaction type, and the energy resolution function of the detector.

@@

Since it is a lot of effort to solve this double integral numerically,
we split up the two integrations. First, we evaluate the integral over
$\hat{E}$, where the only terms containing $\hat{E}$ are
$k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
$ V_\beta(\hat{E}-E')$. We define:
\begin{eqnarray}
\label{eq:e_res} 
R_\beta^{\text{IT}}(E,E')\,\epsilon_\beta^{\text{IT}}(E')
 \equiv
\int\limits_0^\infty d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
\,V_\beta(\hat{E}-E')\,. 
\end{eqnarray}
Thus, $R_\beta^{\text{IT}}(E,E')$ describes the energy response of 
the detector, \ie , a neutrino with a (true) energy $E$ is reconstructed
with an energy between $E'$ and $E'+dE'$ with a probability
$R_\beta^{\text{IT}}(E,E') dE'$. The function $R(E,E')$ is also often called ``energy resolution function''. Actually, its internal representation
in the software is a smearing matrix\index{Energy resolution}\index{Smear matrix}. The function $\epsilon_\beta^{\text{IT}}(E')$ will later be refered to as ``post-smearing efficiencies'', since it will allow us to define cuts and threshold functions {\em after} the smearing is performed, \ie, as function of $E'$. The detailed definition and initialization of the energy resolution function is described in \Sec~\ref{sec:energy}.

Eventually, we can write down the number of events per bin $i$ \index{Bin} and channel $c$ as
\begin{equation}
\label{eq:channel}
n_i^c=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \quad
\frac{dn_{\beta}^{\text{IT}}}{dE'} (E') \,
\end{equation}
where $\Delta E_i$ is the bin size of the $i$th energy bin.
This means that one has to solve the integral
\begin{eqnarray}
\label{eq:events_bin}
n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad \int\limits_0^\infty dE \,\, \Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
R^c(E,E')\,
\epsilon^c(E')\,.
\end{eqnarray} 
Note that the events are binned according to their \emph{reconstructed} energy.

A simple channel definition in \GLOBES\ consists of the flux,
the CP-sign of the initial state, the initial flavor, the final flavor,
the cross sections, and the energy resolution function. In order to refer to
flux, cross sections, and energy resolution functions, they have to be 
defined before with their {\tt \#name} in the respective environments. 
Thus, a simple definition of a channel is
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
>}
\end{quote}

\index{Oscillation!switching off}
It is also possible to define a channel as no-oscillation by using the
prefix \GLB{NOSC\_} in either the initial  flavour or 
the final flavour, like this
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: NOSC\_muon: muon: \#cross: \#energy\\
>}
\end{quote}

 In that case
all diagonal probabilities $P_{\alpha\alpha}$ are unity and all off-diagonal
probablities  $P_{\alpha\beta}$ are zero. This is \eg\ especially useful for
neutral current events since they do not depend on any oscillation 
parameters\footnote{at least in the absence of sterile neutrinos}. The channels
marked as \GLB{NOSC\_} are only computed by a call to \GLB{glbSetRates}
and not recomputed in the subsequent fit (which calls \GLB{glbSetNewRates}).
Therefore this feature can be used to speed up the rate 
computation considerably, especially in cases where a large set
of channels exist which are only used for the computation of backgrounds.
Usually it is an excellent approximation to treat backgrounds as if they
would not be affected by oscillations.

Note that the energy environment will be described in the next section. 
In addition, one can define pre- and post-smearing effects together
with the channels, which will also be introduced together with the
energy resolution function in the next section.

@@

The definition and implementation of the energy resolution function is 
rather sophisticated in \GLOBES . In particular, since this aspect is 
 time-critical, several algorithms are implemented. The choice of the 
proper algorithm depends on the experiment and the frequencies of the
 involved neutrino oscillations.

>>>>

In order to implement a experiment definition including various
sources of systematical errors, we use several abstraction levels. 
The first level is the so-called ``channel'', \index{norm}{Channel}
 which is the link between 
the oscillation physics and the detection properties for a specfific oscillation pattern (\cf, \figu{channel}). A channel specifies the mapping of a specific neutrino flavor produced by the source onto a reconstructed neutrino flavor.
For example, a muon neutrino oscillates into an electron neutrino and subsequently interacts via quasi-elastic charged current scattering. The measured energy and direction of
the secondary electron in the detector then allows to reconstruct the neutrino energy. The connection from the source flux of the muon neutrino, via the  probability to appear as a electron neutrino, to its detection properties (such as cross sections and energy smearing) is encapsulated into the channel.

@@

The channels are the building blocks for the so-called ``rules''.
\index{norm}{Rule} In general, a rule consists of one or more ``signal'' and ``background'' oscillation channels, which are normalized with efficiencies
(\cf, \figu{rule}). The event numbers from these channels are added {\em before} the $\Delta \chi^2$-value is calculated.\footnote{Note that
in this manual, we $\chi^2$ and $\Delta \chi^2$ are equal, since
for simulated data $\Delta \chi^2 = 0$ at the best-fit point. Thus, we
are using $\chi^2$ and $\Delta \chi^2$ as equal quantities.} In addition, each rule implements an {\em independent} systematics, such as signal and background normalization errors. Eventually, each rule gives a $\Delta \chi^2$-value, and the total $\Delta \chi^2$ of one experiment is obtained by adding the $\Delta \chi^2$'s of all rules (\cf, \figu{experiment}). 
 An example for a rule could look like this: We want to detect electron
  neutrino appearance (``signal''), where the overall efficiency for 
  quasi-elastics electron neutrino events is $0.4$. There is a fraction of 
  $0.01$ of all neutral current events which are mis-identified as 
  quasi-elastic electron neutrino events (``background''). The neutral 
  current fraction is only known within $10\%$ (``background uncertainty'') 
  and there is
an energy scale uncertainty of $100\,\mathrm{MeV}$ (``energy calibration error'').
All this systematics is independent of the other rules.  Thus, a rule connects the event rates to the calculation of a $\Delta \chi^2$ which properly includes systematical errors. The resulting $\Delta \chi^2$ is then the starting point for the oscillation physics analysis. Note again that
\begin{itemize}
\item
 Within each rule the event numbers are added.
\item
 Within each rule the systematics is treated independently from the other rules.
\item
 For each rule the $\Delta \chi^2$ is computed; the $\Delta \chi^2$'s from all rules are added.
\end{itemize}

@@

\index{aedl}{NEXT@{\tt \#NEXT\#}}
If you want to have several experiments in one file, separate the different
 experiments by 
\begin{quote}
{\tt    \#NEXT\#}
\end{quote}
This command resets the counters for channels, rules and energy resolution
functions. All variables have their scope limited 
by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}.  This allows 
a consistent treatment of various experiments in one file.

\index{aedl}{inlcude@{\tt include}}
As another feature of AEDL one can use include files with the {\tt include} command. Includes can be nested up to {\tt MAX\_INCLUSION\_DEPTH}, which is currently set to $10$. Error reporting works 
 for nested includes, too. The included file is not required to begin 
 with {\tt \%!GLoBES} to facilitate cut and paste:
\begin{quote}
{\tt include "./file\_1"}
\end{quote}
With this include mechanism, one can use constructions such as 
\begin{quote}
{\tt    include  "NuFact.gls"\\
        \#NEXT\#\\
        include   "JHFHK.gls"
}
\end{quote}
in order to initialize a combined analysis of the experiments defined in the files {\tt NuFact.gls} and {\tt JHFHK.gls}. As one sees from the example above
file names are always quoted in \AEDL.\index{norm}{File names}
Even if one uses the
automatic variable {\tt \#CC} in both experiments, 
but the cross section data are different (for example, because of different target nuclei), the correct 
cross section data will be applied to each of the experiments. 
Note that, alternatively, one can 
also load both files successively.

Furthermore, one can define constants such as
\begin{quote}
{\tt
Pi = 3.141
}
\end{quote}
These constants can not only be defined within one \AEDL\ file but also
by the calling C program, which allows to use a simple but powerful variable
substitution mechanism as described in~\Sec~\ref{sec:aedlparams}.
\index{norm}{External \AEDL\ parameters}

Alos some simple algebraic manipulations are possible, such as
\begin{quote}
{\tt
Pi+1\\
\verb+Pi^2+\\
sin(Pi/2)\\
}
\end{quote}
The following mathematical functions from {\tt <math.h>} are available: 
{\tt sin}, {\tt cos}, {\tt tan}, {\tt asin}, {\tt acos}, {\tt atan}, 
{\tt log}, {\tt log10}, {\tt exp}, {\tt sqrt}.
\index{aedl}{sin@{\tt sin}}
\index{aedl}{cos@{\tt cos}}
\index{aedl}{tan@{\tt tan}}
\index{aedl}{asin@{\tt asin}}
\index{aedl}{acos@{\tt acos}}
\index{aedl}{atan@{\tt atan}}
\index{aedl}{log@{\tt log}}
\index{aedl}{log10@{\tt log10}}
\index{aedl}{exp@{\tt exp}}
\index{aedl}{sqrt@{\tt sqrt}}

These functions can be used everywhere, where
otherwise only a scalar number would appear. However, they can not be
applied to lists, such as {\tt sin(\{1,2,3\})} will not work.

@@

\index{aedl}{target mass@{\tt \$target\_mass}}
The quantity which can be used to scale the overall integrated luminosity of an experiment, is the fiducial detector mass. For example,
\begin{quote}
{\tt \$target\_mass = 50.0 }
\end{quote}
defines a $50 \, \mathrm{kt}$ detector for a neutrino factory.

\index{aedl}{{\tt flux}!time@{\tt "@time}}
There are two principal ways to initialize a neutrino flux: 
Either one can use 
a built-in source, or one can provide a file. In both cases,
a flux is defined by the environment {\tt flux}, such as
\index{aedl}{{\tt flux}}
\begin{quote}
  {\tt flux(\#name)<\\
\tb $\ldots$\\
\tb @time = 8.0 \\
>}
\end{quote}
with a running time of $8$ years. Note that the running time is used within
the {\tt flux} environment. This feature can be used to load the neutrino and antineutrino fluxes separately, in order to combine them with different
running times within one experiment. The name of the flux {\tt \#name} will later be refered to in the channel definitions.

For a built-in neutrino source, one has to specify which
built-in spectrum has to be used, as well as its parameters. The software
will then automatically calculate the neutrino spectrum. Note that in this
case, there is no degree of freedom in the choice of the source units.
Currently, two built-in fluxes are available: $\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}). In these cases, the muon energy (enery of the parent particle) has to be specified together with the number of useful decays  per year. Thus, an example to set up a neutrino factory flux is
\index{aedl}{{\tt flux}!builtin@{\tt "@builtin}}
\index{aedl}{{\tt flux}!parent energy@{\tt "@parent\_energy}}
\index{aedl}{{\tt flux}!stored muons@{\tt "@stored\_muons}}
\begin{quote}
{\tt flux(\#mu\_plus)<\\
\tb  @builtin = 1\\
\tb  @parent\_energy = 50.0\\
\tb  @stored\_muons = 5.33e+20\\
\tb  @time = 8.0\\
>}
\end{quote}
%
\index{aedl}{{\tt flux}!power@{\tt "@power}}
\index{aedl}{{\tt flux}!norm@{\tt "@norm}}
\index{aedl}{{\tt flux}!flux file@{\tt "@flux\_file}}
For a user-defined flux, one has to give it the file name:
\begin{quote}
{\tt flux(\#user)<}\\
{\tt \tb @flux\_file = "user\_file\_1.dat"\\
\tb @time = 2.0\\
\tb @power = 4.0\\
\tb @norm = 1e+8}\\
{\tt >}
\end{quote}
In this case, the {\tt @norm} variable is an overall normalization which defines a conversion factor from the fluxes in the file to the units in \GLOBES . In general, there are many ways to give the source power of a 
neutrino source, such as neutrinos per proton on target per area per time frame. Right now, each flux has its own normalization factor, which is
not always straightfoward to calculate. Often, one has to take into account
many things, such as the number of target particles per unit mass. 
In addition, the fluxes will be rescaled by $1/L^2$, which means that the
normalization must contain a factor $L_0^2$. Here $L_0$ is the distance from the source for which the flux is given to the actual neutrino production region. At the end, it is left to the user to ensure that the 
numbers in the flux file give, after the multiplication with {\tt @norm}, 
the proper numbers of produced neutrinos corresponding to the chosen target power {\tt @power}. Usually this adjustment of {\tt @norm} is performed by comparison with known energy spectra for a specific experiment.

\index{aedl}{{\tt flux}!flux file@{\tt "@flux\_file}}
The software assumes that the given flux file has seven columns and
501 lines with equidistant energies. The format is:
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
In order to access fluxes at arbitrary energies, linear interpolation 
is used by the software. In general, it is advisable to provide the flux between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}), since these values are used by the software. However, if the energy leaves the range of values given in the file, zero is returned. 
Note that unused (and therefore often unknown) fluxes can be initialized with zeros, but can not just be omitted. For example, if one wants to define an experiment with different running times with neutrinos and antineutrinos, one can just use the same flux file in two different {\tt flux} environments with two different running times {\tt @time}. The unused fluxes in each case will then just be ignored.

@@

\index{aedl}{baseline@{\tt \$baseline}}
The baseline\index{norm}{Baseline} is given by
\begin{quote}
{\tt \$baseline = 3000.0 }
\end{quote}
Note that baseline lengths are always assumed to be in
kilometers.

@@

\index{aedl}{profiletype@{\tt \$profiletype}}
\index{aedl}{densitysteps@{\tt \$densitysteps}}
\index{aedl}{lengthtab@{\tt \$lengthtab}}
\index{aedl}{densitytab@{\tt \$densitytab}}

Furthermore, the matter density profile along the baseline
has to be specified. The simplest matter density profile is a constant matter density profile with the average matter density from the PREM\cite{Stacey} onion shell model of the Earth\index{norm}{PREM}\index{norm}{Earth matter density}:
\begin{quote}
{\tt \$profiletype=1 }
\end{quote}
%
For a better approximation of the realistic Earth matter density profile, one can use an arbitrary number of equidistant steps of the PREM profile:
\begin{quote}
{\tt \$profiletype=2 } \\
{\tt \$densitysteps=20 }
\end{quote}
Note that the value of {\tt \$densitysteps} is time-critical, since the
computation time of oscillation probabilities is directly 
proportional to the number of layers.
%
As a  third possibility, one can specify the matter density profile 
manually with a list of thicknesses and densities of the matter density layers. This example uses two density steps with two different densities:
\begin{quote}
{\tt \$profiletype=3 } \\
{\tt \$densitytab=\{2.8, 3.5\}}\\
{\tt \$lengthtab=\{1000.0, 2000.0\}}\\
\end{quote}
It is important that both lists habe the same length and that the  thicknesses given in  {\tt \$lengthtab} add up to the value of
{\tt \$baseline}. In addition, matter densities are always given in $g/cm^3$.
%
This approach can also be used for a constant matter density profile with
a specific matter density:
\begin{quote}
{\tt \$profiletype=3 } \\
{\tt \$densitytab=\{3.5\}}\\
{\tt \$lengthtab=\{3000.0\}}\\
\end{quote}
The possible options for matter density profiles are summarized in \Tab~\ref{tab:profiletypes}.

@@

Cross sections\index{norm}{Cross sections} will later be used as part of the 
channel definition (see \Sec~\ref{sec:channel}). Similar to the source 
fluxes, they are provided by the user as a data file:
\index{aedl}{{\tt cross}}
\index{aedl}{{\tt cross}!cross file@{\tt  "@cross\_file}}
\begin{quote}
{\tt cross(\#name)<}\\
{\tt \tb @cross\_file ="user\_file\_1.dat"}\\
{\tt >}
\end{quote}  
This cross section can later be refered to by {\tt \#name}.

Cross sections are in \GLOBES\ given as differential cross section per energy:
\begin{equation}
\hat\sigma(E)=\sigma(E)/E\,\left[ 10^{-38}\,
\frac{\mathrm{cm}^2}{\mathrm{GeV}^2} \right]
\end{equation}
The software assumes that the cross section files are text files with 
seven columns and $1001$ lines of the form
\index{aedl}{{\tt cross}!cross file@{\tt  "@cross\_file}}
\begin{quotation}
$\mathrm{log}_{10} E\quad
\hat\sigma_{\nu_e}\quad
\hat\sigma_{\nu_\mu}\quad
\hat\sigma_{\nu_\tau}\quad
\hat\sigma_{\bar\nu_e}\quad
\hat\sigma_{\bar\nu_\mu}\quad
\hat\sigma_{\bar\nu_\tau}$
\end{quotation}
Here the logarithms of the energy values have to be equidistant. For 
arbitrary energies, linear interpolation is used. If the energy leaves the
range of values given in the file, $0.0$ will be assumed. In general, it is advisable to provide the cross sections in the range between {\tt \$sampling\_min} and {\tt \$sampling\_max} (\cf, \Sec~\ref{sec:energy}).
Unused cross sections have to be filled with zeros, and can not
 be just omitted.

@@

\index{aedl}{{\tt channel}|(}
Channels\index{norm}{Channel} in \GLOBES\ represent an intermediate level 
between the pure oscillation physics given by the oscillation probability
$P_{\alpha\beta}$, and the total event rates composed of signal and 
background. A channel describes the path from one
initial neutrino flavor in the source to the event rates in the detector for one specific interaction type (IT) and final flavor.  
Therefore, a channel contains the description of the 
initial neutrino flavor, its CP eigenvalue (neutrino or antineutrino), 
the detected neutrino flavor, the interaction cross sections for the chosen interaction type, and the energy resolution function of the detector.

@@

Since it is a lot of effort to solve this double integral numerically,
we split up the two integrations. First, we evaluate the integral over
$\hat{E}$, where the only terms containing $\hat{E}$ are
$k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
$ V_\beta(\hat{E}-E')$. We define:
\begin{eqnarray}
\label{eq:e_res} 
R_\beta^{\text{IT}}(E,E')\,\epsilon_\beta^{\text{IT}}(E')
 \equiv
\int\limits_0^\infty d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
\,V_\beta(\hat{E}-E')\,. 
\end{eqnarray}
Thus, $R_\beta^{\text{IT}}(E,E')$ describes the energy response of 
the detector, \ie , a neutrino with a (true) energy $E$ is reconstructed
with an energy between $E'$ and $E'+dE'$ with a probability
$R_\beta^{\text{IT}}(E,E') dE'$. The function $R(E,E')$ is also often called ``energy resolution function''. Actually, its internal representation
in the software is a smearing matrix\index{norm}{Energy resolution}\index{norm}{Smear matrix}. The function $\epsilon_\beta^{\text{IT}}(E')$ will later be refered to as ``post-smearing efficiencies'', since it will allow us to define cuts and threshold functions {\em after} the smearing is performed, \ie, as function of $E'$. The detailed definition and initialization of the energy resolution function is described in \Sec~\ref{sec:energy}.

Eventually, we can write down the number of events per bin $i$ \index{norm}{Bin} and channel $c$ as
\begin{equation}
\label{eq:channel}
n_i^c=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \quad
\frac{dn_{\beta}^{\text{IT}}}{dE'} (E') \,
\end{equation}
where $\Delta E_i$ is the bin size of the $i$th energy bin.
This means that one has to solve the integral
\begin{eqnarray}
\label{eq:events_bin}
n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad \int\limits_0^\infty dE \,\, \Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
R^c(E,E')\,
\epsilon^c(E')\,.
\end{eqnarray} 
Note that the events are binned according to their \emph{reconstructed} energy.

A simple channel definition in \GLOBES\ consists of the flux,
the CP-sign of the initial state, the initial flavor, the final flavor,
the cross sections, and the energy resolution function. In order to refer to
flux, cross sections, and energy resolution functions, they have to be 
defined before with their {\tt \#name} in the respective environments. 
Thus, a simple definition of a channel\index{aedl}{{\tt channel}} is
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
>}
\end{quote}

\index{norm}{Oscillation!switching off}
It is also possible to define a channel as no-oscillation by using the
prefix {\tt NOSC\_}\index{aedl}{{\tt channel}!{\tt NOSC\_}} 
in either the initial  flavour or 
the final flavour, like this
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: NOSC\_muon: muon: \#cross: \#energy\\
>}
\end{quote}

 In that case
all diagonal probabilities $P_{\alpha\alpha}$ are unity and all off-diagonal
probablities  $P_{\alpha\beta}$ are zero. This is \eg\ especially useful for
neutral current events since they do not depend on any oscillation 
parameters\footnote{at least in the absence of sterile neutrinos}. The channels
marked as {\tt NOSC\_} are only computed by a call to \GLB{SetRates}
and not recomputed in the subsequent fit (which calls \GLB{SetNewRates}).
Therefore this feature can be used to speed up the rate 
computation considerably, especially in cases where a large set
of channels exist which are only used for the computation of backgrounds.
Usually it is an excellent approximation to treat backgrounds as if they
would not be affected by oscillations.

Note that the energy environment will be described in the next section. 
In addition, one can define pre- and post-smearing effects together
with the channels, which will also be introduced together with the
energy resolution function in the next section.
\index{aedl}{{\tt channel}|)}

@@

\index{aedl}{{\tt energy}|(}
The definition and implementation of the energy resolution function is 
rather sophisticated in \GLOBES . In particular, since this aspect is 
 time-critical, several algorithms are implemented. The choice of the 
proper algorithm depends on the experiment and the frequencies of the
 involved neutrino oscillations.
